---
title: "ST 8114 group project"
author: "Brett Farmer, Mark Hill"
date: "9/17/2020"
output:
  pdf_document:
    toc: yes
  html_document:
    dev: svg
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float:
      colllapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# library/read in of data {-}
```{r, message=FALSE}
rm(list = ls())
library(tidyverse)
library(ggpubr)
library(car)
library(broom)
library(knitr)
library(Hmisc)
library(patchwork)
library(mosaic)
```

```{r, warning= FALSE}

dat <- read.csv("cohort_DIABP_female.txt", header = TRUE, sep = "")
dat$SMOKING <- as.factor(dat$SMOKING)
dat$HTNMED <- as.factor(dat$HTNMED)
attach(dat)
describe(dat)
```

# Test some model assumptions first
```{r}
densityPlot(DIABP)

ggplot(dat)+
  stat_density(aes(x= DIABP))

ggplot(dat)+
  stat_density(aes(x= DIABP, fill= HTNMED) )
```
Here we plotted density of the HTNMED groups because the sample of HTNMED == "yes" is less than the threshold for large sample sizes (n=17) 

```{r}
ggplot(dat, aes(sample = DIABP))+
  stat_qq()+
  stat_qq_line()

qqPlot(DIABP)

shapiro.test(DIABP)

```
It would appear from the density and qqplots that the distribution of the response variable is slightly skewed to the right. The very small p-value in the test would indicate we would reject the null hypothesis that our diastolic blood pressure values come from a normal distribution. Though, with a large sample size (n= 200) the `shapiro.test` detects even a very tiny departure from normality. For our case this departure will not impact our `t.test` result in any meaningful way because it is robust against this violation when used with very large sample sizes.

# How is each explanatory variable associated with the response variable? <br> (DIABP ~ c(AGE, SMOKING, HTNMED))

## EDA: model with all three variables

```{r}
model3 <- lm(DIABP ~ SMOKING+ AGE + HTNMED)
summary(model3)$coef %>% kable()
```


## DIABP ~ AGE
```{r}

cor(AGE, DIABP)
model1 <- lm(DIABP ~ AGE)
summary(model1)$coef %>% kable()
```

```{r, message=FALSE}
ggplot(dat,
       aes(x = AGE, y= DIABP))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_pubclean()
```

## DIABP ~ SMOKING
```{r}
tests <- list(c("current-smoker", "non-smoker"), 
              c("current-smoker", "past-smoker"), 
              c("non-smoker","past-smoker"))
ggplot(dat, aes(y= DIABP, x = SMOKING, fill= SMOKING))+
  geom_boxplot()+
  stat_compare_means(comparisons = tests, method = "t.test")+
  ggtitle("Comparison of Diastolic Blood Pressure distribution for factor \n SMOKING")+
  theme_pubclean()+
  theme(legend.position = "none")

compare_means(DIABP ~ SMOKING, data = dat, method = "t.test") %>% kable()

model2 <- lm(DIABP ~ SMOKING)
summary(model2)$coef %>% kable()
```

## DIABP ~ HTNMED
```{r}
ggplot(dat,
      aes(x= HTNMED, y= DIABP, fill= HTNMED))+
    geom_boxplot()+
    stat_compare_means(method = "t.test", method.args = list(var.equal = TRUE, alternative= "greater"))+
    ggtitle("Comparison of Diastolic Blood Pressure distribution for factor \n HYPERTENSION MEDS")+
  theme_pubclean()+
  theme(legend.position = "none")

model3 <- lm(DIABP ~ HTNMED)
summary(model3)$coef %>% kable()
```
In a linear model, the very small p-value [2,4] for the yes category suggests there is a very strong relationship between the use of hypertension medication and diastolic blood pressure.

```{r}
t.test(DIABP ~ HTNMED, alternative = "less", var.equal =  TRUE)

```
Here our `t.test` suggests that we can move to reject the null hypothesis and conclude that there is a statistically significant differnce between these groups.

## Model which contains both categorical predictor vars

```{r}
model4 <- lm(DIABP ~ SMOKING + HTNMED)
Anova(model4) %>% kable()
```

```{r, results='asis'}
summary(model4)$coef %>% kable()
```

```{r}
bp <- ggplot(dat)+
  geom_boxplot(aes(y= DIABP))+
  theme_pubclean()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
ag <- ggplot(dat)+
  geom_boxplot(aes(y= AGE))+
  theme_pubclean()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
bp+ag

```


```{r}
sumstats <- as.data.frame(favstats(dat[dat$HTNMED == "yes", 1]))
sumstats[2,] <- favstats(dat[dat$HTNMED == "no", 1])
sumstats$group <- c("yes", "no")
sumstats <- sumstats[,c(10,1:8)]
sumstats %>% kable(row.names = FALSE, format = "latex", booktabs= TRUE)
```

