---
title: |
      Lab Homework Module:5   
      ST8114  
      mjh100  
      
author: "Mark Hill"
date: "10/30/2020"
output:
  pdf_document: 
    toc: true
    toc_depth: 2
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(patchwork)
library(nortest)
library(broom)
library(purrr)
```


\newpage

# Practice 1
```{r}
estate <- read.csv("~/ST 8114/lab/module 5/data/estate.txt", sep = " ", header = TRUE)
```

## a
```{r, fig.height=4}

ggplot(estate, aes(x= finished, y= price))+
  geom_point()+
  geom_smooth(method = "lm", se=F)
```

\newpage

## b
```{r}
model <- lm(price ~ finished, estate)
summary(model)

cat("yhat=", model$coefficients[1], "+", model$coefficients[2],"x")
```
We would expect the cost of houses to increase by 0.1526 * thousand dollars for each additional square foot in size. ie. $152 more for each additional square foot.

\newpage

# Practice 2
```{r}
garbage <- read.csv("~/ST 8114/lab/module 5/data/garbage.txt", header = TRUE)
```

## a

```{r, fig.height=4.5}
a <- ggplot(garbage, aes(x= volume))+
      geom_boxplot()

b <- ggplot(garbage, aes(sample= volume))+
  stat_qq()+
  stat_qq_line()

a / b
```
These appear to be normally distributed values

\newpage

## b

```{r}
t1 <- shapiro.test(garbage$volume)
t2 <- pearson.test(garbage$volume)
t3 <- sf.test(garbage$volume)
t4 <- ad.test(garbage$volume)
tab <- map_df(list(t1,t2,t3,t4),tidy)
tab
```
No rejection of $H_0$ in any of the tests.

## c

```{r}
t5 <- t.test(garbage$volume, conf.level = .9)
cat("90% confidence interval for mean of garbage$volume",t5$conf.int)                
```

## d

```{r}
t6 <- t.test(garbage$volume, conf.level = .90, mu= 1600, alternative = "greater")
t6
```
There is evidence, at $\alpha$ = 0.1, to reject $H_0: \mu \leq 1600$. Thus the two week schedule is desirable since they would usually pick up more than 1600 cubic feet.

\newpage

# Practice 3
```{r}
wgtgain <- read.csv("~/ST 8114/lab/module 5/data/wgtgain.txt", header = TRUE, sep = " ")
```

## a

```{r}
wgtgain$diff <- wgtgain$after - wgtgain$before
```
## b

```{r, fig.height= 3}
ggplot(wgtgain, aes(sample= diff))+
  stat_qq()+
  stat_qq_line()
```

## c

```{r}
t1 <- tidy(shapiro.test(wgtgain$diff))
t1
```

## d

```{r}
t2 <- t.test(x= wgtgain$after, y= wgtgain$before, paired = TRUE)
cat("95% confidence interval for weight gain after treatment is", "(",t2$conf.int,") kg.")
```

## e

```{r}
t3 <- t.test(wgtgain$diff, alternative = "two.sided", mu= 7)
```

## f

```{r}
t4 <- t.test(wgtgain$diff, alternative = "greater", mu= 7)
```

## g

```{r}
t5 <- t.test(wgtgain$diff, alternative = "less", mu= 7)

tab <- map_df(list(t3,t4,t5), tidy)
tab
```
  1. p-value < $\alpha$ =0.05, accept $H_a$ that differences are not equal to 7 Kg.
  2. p-value $\approx$ 1, fail to reject $H_0$ that weight gain after the treatment is greater than 7 Kg.
  3. p-value < $\alpha$ =0.05, accept $H_a$ that differences are less than 7 Kg.

\newpage

# Practice 4
```{r}
bonuses <- read.csv("~/ST 8114/lab/module 5/data/bonuses.txt", header = TRUE, sep = " ",
                    stringsAsFactors = TRUE)
```

## a
```{r, results='hold', fig.height=3}
ggplot(bonuses)+
  geom_boxplot(aes(y= bonus))+
  facet_wrap(~gender)

ggplot(bonuses,aes(sample= bonus))+
  stat_qq()+
  stat_qq_line()+
  facet_wrap(~gender)
```

\newpage

## b
```{r}
do.call("rbind",
        with(bonuses, tapply(bonus, gender,
              function(x) 
                unlist(shapiro.test(x)
                       [c("statistic", "p.value")])
                      )
             )
        )
```
Based on the Q-Q plot, boxplot, and the small p-value, the male sample follows a non-normal distribution. It exhibits a pattern of right skewness.

## c

```{r}
t.test(bonuses$bonus ~ bonuses$gender)
```
Here, at the assumed $\alpha$ = 0.05 we would fail to reject $H_0  : \mu_1 = \mu_2$  
<b><b><b>  
But, our assumption of normality has been violated from one of the groups. Lets see if the findings will be different utilizing a nonparametric test.

```{r}
wilcox.test(bonuses$bonus ~ bonuses$gender, correct= FALSE)
```
Again, we don't reject the null in this `two.sided` test. This less "efficient" test leaves us even farther away despite the aforementioned normality violation.

\newpage

## d

```{r}
var.test(bonuses$bonus~ bonuses$gender)
```


## e

```{r}
t.test(bonuses$bonus ~ bonuses$gender, var.equal= TRUE, alternative= "less")
```
Assuming the groups have equal variance and specifying the alternative presents a different result in our `t.test`. The last test showed a difference in means hence the direction of the alternative. Even a two tailed test results in p-value  $<\alpha = 0.05$. Conclusion: accept $H_a$ : female bonuses are less than male bonuses.
